---
title: "NumericColumns Details"
author: "Rony Sulca"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

The package `numericColumns` is built to ease the selection of numeric columns, or columns with all elements
having numeric equivalents, from datasets in data frame, data table, or CSV formats. This is meant to shorten
the time it takes to subset datasets to their numerical columns in order to quickly apply numerical-based models.


## check_numeric_object

`check_numeric_object` asseses if an `object` can be a numeric value. It returns `TRUE`
if the `object` is `numeric` or if it detects numeric values in character form,
such as "1", "3.4", "-76", etc.
If `with_bool` is set to TRUE, `check_numeric_object` assumes booleans
can also have an equivalent numeric form (being 1 and 0).

### Examples:

Any number, even in string form, will pass the test. Objects like factors, string words, or Booleans will fail.

```{r, results='asis', eval=FALSE}
check_numeric_object(1) # [1] True
check_numeric_object("-2.7") # [1] TRUE
check_numeric_object("bacon1") # [1] FALSE
```

With the `with_bool` argument set to TRUE, booleans will also pass the test.

```{r, results='asis', eval=FALSE}
check_numeric_object(TRUE) # [1] False
check_numeric_object(TRUE, with_bool = TRUE) # [1] TRUE
```

## check_numeric_list

`check_numeric_list` checks if all of the elements in the `lst` can be a numeric value. It
returns `TRUE` if all of the elements of `lst` are `numeric` or are numbers in 
character form, such as "1", "3.4", "-76", etc.
If `with_bool` is set to TRUE, `check_numeric_list` assumes booleans
can also have an equivalent numeric form (being 1 and 0), and returns TRUE if it detects a boolean `lst`.

### Examples

`check_numeric_list` will return `TRUE` if all of the elements of `lst` are numeric or have a numeric equivalent.
If there exists at least one element who fails the test (such as a word, factor, or boolean), the function will return `FALSE`.

```{r, results='asis', eval=FALSE}
check_numeric_list(c(1,2,-5,6.7)) # [1] TRUE
check_numeric_list(c("-2.7", 7, 9, -5, "9.8")) # [1] TRUE
check_numeric_list(list("-2.7", 7, 9, -5, "9.8")) # [1] TRUE
check_numeric_list(c("1","3", -5, "bacon")) # [1] FALSE
```

As with `check_numeric_object`, `with_bool` allows for boolean vector or lists to pass the test. Note: We refer
to boolean vectors here because, when booleans are part of an atomic list with other numbers, they will be coerced
to their numeric equivalent, similarly if the atomic list is a character vector.

```{r, results='asis', eval=FALSE}
check_numeric_list(c(TRUE,TRUE, FALSE)) # [1] FALSE
check_numeric_list(c(TRUE,TRUE, FALSE),with_bool = T) # [1] TRUE
```

## extract_numeric_columns

This function returns all of the columns of the given `dataset` whose values are numeric or have a
meaningful numerical equivalent. If `isCSV` is `TRUE`, then `dataset` is taken as a string
depicting the name of a CSV file. `with_bool` specifies if booleans are considered numeric.

The `column` argument is given to bypass which arguments the test is applied to. Columns that fail the test will
still be returned with the rest, but with their entire values converted to `NA`. By default, this argument is 
`FALSE`, the test is applied to all the columns, and only the passing ones are returned.

The following arguments only apply when `isCSV` is `TRUE`: `row_names` specifies if the CSV file must be read with row names. It can take as values `TRUE`, which will set the first column as row names, or a numerical value, which will be selected as the column index for the row names. Note: the column selected as row names must not include duplicates.

An additional `header` argument is available to read CSV files with their headers. Default is `TRUE`.

### Examples:

```{r, results='asis', eval=FALSE}
extract_numeric_columns(mtcars)
extract_numeric_columns("test_data_1.csv", isCSV = TRUE)
extract_numeric_columns("test_data_2.csv", isCSV = TRUE, row_names = TRUE, with_bool = TRUE)
```

The following will return the columns "mpg", and "cyl" which pass the test, as well as the column
"names" which will be coerced to NA's for failing the numeric test.
```{r, results='asis', eval=FALSE}
temp = mtcars
temp$names = row.names(temp)

extract_numeric_columns(temp, columns = c("mpg", "cyl", "names"))
```
